### 2024-05-18 content/chapter_3.tex
:construction: Add viva slides
:triangular_flag_on_post: Viva
:triangular_flag_on_post: Minor amendment 240518

\begin{section}{System Model}
	\begin{figure}[H]
		\centering
		\def\svgwidth{0.9\columnwidth}
		\input{assets/chapter_3/system.eps_tex}
		\caption{An \gls{ris}-aided multi-carrier \gls{miso} \gls{swipt} system.}
		\label{fg:system}
	\end{figure}

@@

	\begin{subsection}{Transmitted Signal}
		Following \cite{Clerckx2018b}, we superpose a multi-carrier modulated information-bearing waveform to a multi-carrier unmodulated power-dedicated deterministic multisine to boost the spectrum and energy efficiency. The information signal transmitted over subband $n \in \mathcal{N} \triangleq \{1, \dots, N\}$ at time $t$ is
		\begin{equation}
			\mathbf{x}_{\mathrm{I},n}(t) = \Re\left\{\mathbf{w}_{\mathrm{I},n} \tilde{x}_{\mathrm{I},n}(t) e^{\jmath 2{\pi}{f_n}{t}}\right\},
		\end{equation}
		where $\mathbf{w}_{\mathrm{I},n} \in \mathbb{C}^{M \times 1}$ is the information precoder at subband $n$, $\tilde{x}_{\mathrm{I},n}\sim\mathcal{CN}(0,1)$ is the information symbol at subband $n$, and $f_n$ is the frequency of subband $n$. On the other hand, the power signal transmitted over subband $n$ at time $t$ is
		\begin{equation}
			\mathbf{x}_{\mathrm{P},n}(t) = \Re\left\{\mathbf{w}_{\mathrm{P},n} e^{\jmath 2{\pi}{f_n}{t}}\right\},
		\end{equation}
		where $\mathbf{w}_{\mathrm{P},n} \in \mathbb{C}^{M \times 1}$ is the power precoder at subband $n$. Therefore, the superposed signal transmitted over all subbands at time $t$ is
		\begin{equation}
			\mathbf{x}(t) = \Re{\left\{\sum_{n=1}^N(\mathbf{w}_{\mathrm{I},n}\tilde{x}_{\mathrm{I},n}(t)+\mathbf{w}_{\mathrm{P},n}){e^{\jmath 2{\pi}{f_n}{t}}}\right\}}.
		\end{equation}
		We also define $\mathbf{w}_{\mathrm{I/P}} \triangleq [\mathbf{w}_{\mathrm{I/P},1}^\mathsf{T},\dots,\mathbf{w}_{\mathrm{I/P},N}^\mathsf{T}]^\mathsf{T} \in \mathbb{C}^{MN \times 1}$.
	\end{subsection}

@@

	\begin{subsection}{Receiving Modes}
		Following the discussion in Section \ref{sc:receiver_architectures}, we investigate the performance of \gls{ts} and \gls{ps} receivers in this work. The \gls{ts} receiver divides each transmission block into orthogonal data and energy sessions with duration $1-\eta$ and $\eta$, respectively. During each session, the transmitter optimizes the waveform for either \gls{wit} or \gls{wpt}, while the receiver activates the information decoder or the energy harvester correspondingly. The duration ratio $\eta$ thus controls the \gls{r-e} trade-off and is independent from the waveform and beamforming design. On the other hand, the \gls{ps} receiver splits the incoming signal into individual data and energy streams with power ratio $1-\rho$ and $\rho$, respectively. The data stream is fed into the information decoder while the energy stream is fed into the energy harvester. During each transmission block, the superposed waveform and splitting ratio are jointly designed to achieve different performance trade-offs.
		Since the \gls{r-e} region of the \gls{ts} receiver can be inferred from that of the \gls{ps} receiver, we focus on the optimization with the \gls{ps} receiver in the following context.
	\end{subsection}

	\begin{subsection}{Information Decoder}
		A major benefit of the superposed waveform is that the multisine is deterministic and its impact on \gls{wit} can be completely eliminated by waveform cancellation or translated demodulation \cite{Clerckx2018b}.
		Therefore, the achievable rate is
		\begin{equation}\label{eq:R}
			R(\boldsymbol{\theta},\mathbf{w}_{\mathrm{I}},\rho) = \sum_{n=1}^N{\log_2\left(1+\frac{(1-\rho)\lvert \mathbf{h}_{n}^\mathsf{H}\mathbf{w}_{\mathrm{I},n} \rvert^2}{\sigma_n^2}\right)},
		\end{equation}
		where $\sigma_n^2$ is the variance of the total noise (at the \gls{rf}-band and during the \gls{rf}-to-baseband conversion) on tone $n$.
	\end{subsection}

@@

			& = \mathbf{w}_{a\mathrm{I},n}^\mathsf{H}\mathbf{h}_n\mathbf{h}_n^\mathsf{H}\mathbf{w}_{\mathrm{I},n}\nonumber\\

@@

	To evaluate the proposed \gls{ris}-aided \gls{swipt} system, we consider the layout in Fig.~\ref{fg:layout} where the \gls{ris} moves along a line parallel to the \gls{ap}-\gls{ue} path. Let $d_{\mathrm{H}}$, $d_{\mathrm{V}}$ be the horizontal and vertical distances from the \gls{ap} to the \gls{ris}, and denote respectively $d_{\mathrm{D}}$, $d_{\mathrm{F}}=\sqrt{d_{\mathrm{H}}^2+d_{\mathrm{V}}^2}$, $d_{\mathrm{B}}=\sqrt{(d_{\mathrm{D}}-d_{\mathrm{H}})^2+d_{\mathrm{V}}^2}$ as the distance of direct, forward and backward links. Set $d_{\mathrm{D}}=\qty{12}{\meter}$ and $d_{\mathrm{H}}=d_{\mathrm{V}}=\qty{2}{\meter}$ as reference. The path loss of direct, forward and backward links are denoted by $\Lambda_{\mathrm{D}}$, $\Lambda_{\mathrm{F}}$ and $\Lambda_{\mathrm{B}}$, respectively. We consider a large open space Wi-Fi-like environment at center frequency \qty{2.4}{\GHz} where the channel follows IEEE TGn channel model D \cite{Erceg2004}. Specifically, the path loss exponent is \num{2} (i.e., free-space model) up to \qty{10}{\meter}, and \num{3.5} onwards to penalize the potential blockage. All fadings are modeled as \gls{nlos} with tap delays and powers specified in model D, and the tap gains are modeled as i.i.d. \gls{cscg} variables. Rectenna parameters are set to $k_2=0.0034$, $k_4=0.3829$, $R_{\mathrm{A}}=\qty{50}{\ohm}$ \cite{Clerckx2016a} such that $\beta_2=0.17$ and $\beta_4=957.25$. We also choose the average \gls{eirp} as $P=\qty{40}{dBm}$, the receive antenna gain as \qty{3}{dBi}, the scaling ratio as $\alpha=2$, and the tolerance as $\epsilon=10^{-8}$. To further reduce the complexity, we assume $\delta=\rho$ and perform a one-dimensional search for the \gls{lc}-\gls{bcd} algorithm. Each \gls{r-e} point is averaged over \num{200} channel realizations, and the $x$-axis is normalized to per-subband rate $R/N$.

@@

		Fig.~\ref{fg:channel_amplitude} reveals how \gls{ris} influences the sorted equivalent subchannel amplitude for one channel realization. Due to the flexible subchannel design enabled by passive beamforming, the optimal amplitude distribution for \gls{wit} and \gls{wpt} are dissimilar. Under the specified configuration, the \gls{wpt}-optimized \gls{ris} aligns the strong subbands to exploit the rectifier nonlinearity. On the other hand, the \gls{wit}-optimized \gls{ris} provides a fair gain over all subchannels when $L$ is sufficiently large. This is reminiscent of the \gls{wf} scheme at high \gls{snr}, but is realized by channel alignment by \gls{ris} instead of resource allocation by transmitter.
	\end{subsection}

@@

			The impacts of the number of transmit antennas $M$ and the \gls{ris} elements $L$ on the \gls{r-e} behavior are revealed in Figs.~\subref*{fg:re_tx} and \subref*{fg:re_reflector}. First, it is observed that adding either active or passive elements can improve the equivalent \gls{snr}, which produces a nearly concave \gls{r-e} region and favors the \gls{ps} receiver. Second, the conventional \gls{leh} model leads to a power-inefficient design. To investigate the performance loss, we truncate the \gls{dc} objective function \eqref{eq:z} at $n_0=2$ such that (i) in the passive beamforming problem, $z(\bar{\mathbf{\Theta}}) = {\beta_2}{\rho}(t_{\mathrm{I},0}+t_{\mathrm{P},0})/2$ and no \gls{sca} is required; (ii) in the waveform design problem, the \gls{wpt}-optimal strategy is the adaptive single sinewave that allocates all power to the multisine at the strongest subband \cite{Clerckx2016a}. As shown in Figs.~\subref*{fg:scaling_tx} and \subref*{fg:scaling_reflector}, those conventional designs do not exploit the harvester nonlinearity and end up with a nearly \qty{20}{dBA} gap compared to the nonlinear model-based \gls{smf} and \gls{gp} designs. Third, doubling $M$ brings a \qty{3}{\dB} gain at the output \gls{snr} and a \qty{12}{dBA} increase at the harvested \gls{dc}, which verified that active beamforming has an array gain of $M$ \cite{Tse2005} with power scaling order $M^2$ under the truncated nonlinear harvester model \cite{Clerckx2016a,Clerckx2018b}. Fourth, when the \gls{ris} is very close to the \gls{ap} or \gls{ue}, doubling $L$ can bring a \qty{6}{\dB} gain at the output \gls{snr} and a \qty{24}{dBA} increase at the harvested \gls{dc}. From the perspective of \gls{wit}, it suggests that passive beamforming can reach an array gain of $L^2$, as indicated by \cite{Wu2019}. An interpretation is that the \gls{ris} coherently combines the incoming signal with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$. From the perspective of \gls{wpt}, it suggests that passive beamforming comes with a power scaling order $L^4$ under the truncated nonlinear harvester model. We then verify this novel observation in a simplified case where the power is uniformly allocated over multisine, all channels are frequency-flat, and $L$ is sufficiently large such that the direct channel becomes negligible. Let $X$ be the cascaded small-scale fading coefficient. The \gls{dc} in such case reduces to
			\begin{equation}
				z = \beta_2 \Lambda_{\mathrm{B}}^2 \Lambda_{\mathrm{F}}^2 \lvert X \rvert^2 L^2 P + \beta_4 \frac{2N^2 + 1}{2N} \Lambda_{\mathrm{B}}^4 \Lambda_{\mathrm{F}}^4 \lvert X \rvert^4 L^4 P^2,
			\end{equation}
			which scales quartically with $L$. Compared with active antennas, \gls{ris} elements achieve higher array gain and power scaling order, but a very large $L$ is required to compensate the double fading of the auxiliary link. These observations demonstrate the \gls{r-e} benefit of passive beamforming and emphasize the importance of accounting for the harvester nonlinearity in the waveform and beamforming design.
		\end{subsubsection}

>>>>

\begin{section}{System Model}
	\begin{figure}[H]
		\centering
		\def\svgwidth{0.7\columnwidth}
		\input{assets/chapter_3/system.eps_tex}
		\caption{A \gls{ris}-aided multi-carrier \gls{miso} \gls{swipt} system.}
		\label{fg:system}
	\end{figure}

@@

	\begin{subsection}{Transmitted Signal}
		Following \cite{Clerckx2018b}, we superpose a multi-carrier modulated information-bearing waveform to an unmodulated power-dedicated multisine to boost the spectrum and energy efficiency.
		It is worth mentioning that the latter is beneficial for \gls{wpt} due to its higher \gls{papr} (as discussed in Section \ref{sc:operation_regions}) and better high-order statistics on expectation (c.f. \eqref{eq:y_I4} and \eqref{eq:y_P4} in Section \ref{sc:energy_harvester}).
		The information signal transmitted over subband $n \in \mathcal{N} \triangleq \{1, \dots, N\}$ at time $t$ is
		\begin{equation}
			\mathbf{x}_{\mathrm{I},n}(t) = \Re\left\{\mathbf{w}_{\mathrm{I},n} \tilde{x}_{\mathrm{I},n}(t) e^{\jmath 2{\pi}{f_n}{t}}\right\},
		\end{equation}
		where $\mathbf{w}_{\mathrm{I},n} \in \mathbb{C}^{M \times 1}$ is the information precoder at subband $n$, $\tilde{x}_{\mathrm{I},n}\sim\mathcal{CN}(0,1)$ is the information symbol at subband $n$, and $f_n$ is the frequency of subband $n$. On the other hand, the power signal transmitted over subband $n$ at time $t$ is
		\begin{equation}
			\mathbf{x}_{\mathrm{P},n}(t) = \Re\left\{\mathbf{w}_{\mathrm{P},n} e^{\jmath 2{\pi}{f_n}{t}}\right\},
		\end{equation}
		where $\mathbf{w}_{\mathrm{P},n} \in \mathbb{C}^{M \times 1}$ is the power precoder at subband $n$. Therefore, the superposed signal transmitted over all subbands at time $t$ is
		\begin{equation}
			\mathbf{x}(t) = \Re{\left\{\sum_{n=1}^N(\mathbf{w}_{\mathrm{I},n}\tilde{x}_{\mathrm{I},n}(t)+\mathbf{w}_{\mathrm{P},n}){e^{\jmath 2{\pi}{f_n}{t}}}\right\}}.
		\end{equation}
		We also define $\mathbf{w}_{\mathrm{I/P}} \triangleq [\mathbf{w}_{\mathrm{I/P},1}^\mathsf{T},\dots,\mathbf{w}_{\mathrm{I/P},N}^\mathsf{T}]^\mathsf{T} \in \mathbb{C}^{MN \times 1}$.
	\end{subsection}

@@

	\begin{subsection}{Receiving Modes}
		Following the discussion in Section \ref{sc:receiver_architectures}, we investigate the performance of \gls{ts} and \gls{ps} receivers in this work.\footnote{It is worth mentioning that many other \gls{swipt} transceiving strategies have been proposed in the literature. In particular, frequency selection is more practical where multisine is used over $n$ subbands and information is carried over the remaining $N-n$ subbands, where $n$ is a design parameter. This can be viewed as a special case of \gls{ps} and its \gls{r-e} region is strictly contained with that of the latter, such that we don't include a special study in this work.} The \gls{ts} receiver divides each transmission block into orthogonal data and energy sessions with duration $1-\eta$ and $\eta$, respectively. During each session, the transmitter optimizes the waveform for either \gls{wit} or \gls{wpt}, while the receiver activates the information decoder or the energy harvester correspondingly. The duration ratio $\eta$ thus controls the \gls{r-e} trade-off and is independent from the waveform and beamforming design. On the other hand, the \gls{ps} receiver splits the incoming signal into individual data and energy streams with power ratio $1-\rho$ and $\rho$, respectively. The data stream is fed into the information decoder while the energy stream is fed into the energy harvester. During each transmission block, the superposed waveform and splitting ratio are jointly designed to achieve different performance trade-offs.
		Since the \gls{r-e} region of the \gls{ts} receiver can be inferred from that of the \gls{ps} receiver, we focus on the optimization with the \gls{ps} receiver in the following context.
	\end{subsection}

	\begin{subsection}{Information Decoder}
		A major benefit of the superposed waveform is that the multisine is deterministic and its impact on \gls{wit} can be completely eliminated by waveform cancellation or translated demodulation \cite{Clerckx2018b}.
		The former is achieved by subtracting the multisine from the received signal at additional signal processing cost, while the latter requires a demodulator with higher saturation power level.
		% , while the latter is achieved by multiplying the received signal with the conjugate of the multisine.
		The achievable rate is thus
		\begin{equation}\label{eq:R}
			R(\boldsymbol{\theta},\mathbf{w}_{\mathrm{I}},\rho) = \sum_{n=1}^N{\log_2\left(1+\frac{(1-\rho)\lvert \mathbf{h}_{n}^\mathsf{H}\mathbf{w}_{\mathrm{I},n} \rvert^2}{\sigma_n^2}\right)},
		\end{equation}
		where $\sigma_n^2$ is the variance of the total noise (at the \gls{rf}-band and during the \gls{rf}-to-baseband conversion) on tone $n$.
	\end{subsection}

@@

			& = \mathbf{w}_{\mathrm{I},n}^\mathsf{H}\mathbf{h}_n\mathbf{h}_n^\mathsf{H}\mathbf{w}_{\mathrm{I},n}\nonumber\\

@@

	To evaluate the proposed \gls{ris}-aided \gls{swipt} system, we consider the layout in Fig.~\ref{fg:layout} where the \gls{ris} moves along a line parallel to the \gls{ap}-\gls{ue} path. Let $d_{\mathrm{H}}$, $d_{\mathrm{V}}$ be the horizontal and vertical distances from the \gls{ap} to the \gls{ris}, and denote respectively $d_{\mathrm{D}}$, $d_{\mathrm{F}}=\sqrt{d_{\mathrm{H}}^2+d_{\mathrm{V}}^2}$, $d_{\mathrm{B}}=\sqrt{(d_{\mathrm{D}}-d_{\mathrm{H}})^2+d_{\mathrm{V}}^2}$ as the distance of direct, forward and backward links. Set $d_{\mathrm{D}}=\qty{12}{\meter}$ and $d_{\mathrm{H}}=d_{\mathrm{V}}=\qty{2}{\meter}$ as reference. The path loss of direct, forward and backward links are denoted by $\Lambda_{\mathrm{D}}$, $\Lambda_{\mathrm{F}}$ and $\Lambda_{\mathrm{B}}$, respectively. We consider a Wi-Fi-like environment at center frequency \qty{2.4}{\GHz} where the channel follows IEEE TGn channel model D \cite{Erceg2004}. To simulate an indoor environment where two rooms are separated by a wall and a \gls{ris} door, we set the path loss exponent to \num{2} up to \qty{10}{\meter} and \num{3.5} onwards to penalize the penetration loss. All fadings are modeled as \gls{nlos} with tap delays and powers specified in model D, and the tap gains are modeled as i.i.d. \gls{cscg} variables. Rectenna parameters are set to $k_2=0.0034$, $k_4=0.3829$, $R_{\mathrm{A}}=\qty{50}{\ohm}$ \cite{Clerckx2016a} such that $\beta_2=0.17$ and $\beta_4=957.25$. We also choose the average \gls{eirp} as $P=\qty{40}{dBm}$\footnote{One examiner has kindly pointed out that the regulation for EIRP is 36 dBm in the 2.4 GHz band. Unfortunately, re-running all the simulations is too time-consuming at this stage and reducing the EIRP by 4dB will only affect the scale of the simulation results without altering any conclusion. We have published the source code on \href{https://github.com/snowztail/irs-aided-swipt-joint-waveform-active-and-passive-beamforming-design-under-nonlinear-harvester-model}{this link} for everyone to verify the results.}, the receive antenna gain as \qty{3}{dBi}, the scaling ratio as $\alpha=2$, and the tolerance as $\epsilon=10^{-8}$. To further reduce the complexity, we assume $\delta=\rho$ and perform a one-dimensional search for the \gls{lc}-\gls{bcd} algorithm. Each \gls{r-e} point is averaged over \num{200} channel realizations, and the $x$-axis is normalized to per-subband rate $R/N$.

@@

		Fig.~\ref{fg:channel_amplitude} reveals how \gls{ris} influences the sorted equivalent subchannel amplitude for one channel realization.\footnote{The results of \gls{ris} with random phase has not been included in Fig.~\ref{fg:channel_amplitude}, since its impact on subchannel amplitude gets averaged out and the result coincides with no \gls{ris}. This can be observed later in Figs. \subref*{fg:re_irs_1mhz} and \subref*{fg:re_irs_10mhz}. In conclusion, the subchannel amplitude gain origins from the optimization of \gls{ris} rather than the extra propagation path.} Due to the flexible subchannel design enabled by passive beamforming, the optimal amplitude distribution for \gls{wit} and \gls{wpt} are dissimilar. Under the specified configuration, the \gls{wpt}-optimized \gls{ris} aligns the strong subbands to exploit the rectifier nonlinearity. On the other hand, the \gls{wit}-optimized \gls{ris} provides a fair gain over all subchannels when $L$ is sufficiently large. This is reminiscent of the \gls{wf} scheme at high \gls{snr}, but is realized by channel alignment by \gls{ris} instead of resource allocation by transmitter.
	\end{subsection}

@@

			The impacts of the number of transmit antennas $M$ and the \gls{ris} elements $L$ on the \gls{r-e} behavior are revealed in Figs.~\subref*{fg:re_tx} and \subref*{fg:re_reflector}. First, it is observed that adding either active or passive elements can improve the equivalent \gls{snr}, which produces a nearly concave \gls{r-e} region and favors the \gls{ps} receiver. Second, the conventional \gls{leh} model leads to a power-inefficient design. To investigate the performance loss, we truncate the \gls{dc} objective function \eqref{eq:z} at $n_0=2$ such that (i) in the passive beamforming problem, $z(\bar{\mathbf{\Theta}}) = {\beta_2}{\rho}(t_{\mathrm{I},0}+t_{\mathrm{P},0})/2$ and no \gls{sca} is required; (ii) in the waveform design problem, the \gls{wpt}-optimal strategy is the adaptive single sinewave that allocates all power to the multisine at the strongest subband \cite{Clerckx2016a}. As shown in Figs.~\subref*{fg:scaling_tx} and \subref*{fg:scaling_reflector}, those conventional designs do not exploit the harvester nonlinearity and end up with a nearly \qty{20}{dB} gap compared to the nonlinear model-based \gls{smf} and \gls{gp} designs. Third, doubling $M$ brings a \qty{3}{\dB} gain at the output \gls{snr} and a \qty{12}{dB} increase at the harvested \gls{dc}, which verified that active beamforming has an array gain of $M$ \cite{Tse2005} with power scaling order $M^2$ under the truncated nonlinear harvester model \cite{Clerckx2016a,Clerckx2018b}. Fourth, when the \gls{ris} is very close to the \gls{ap} or \gls{ue}, doubling $L$ can bring a \qty{6}{\dB} gain at the output \gls{snr} and a \qty{24}{dB} increase at the harvested \gls{dc}. From the perspective of \gls{wit}, it suggests that passive beamforming can reach an array gain of $L^2$, as indicated by \cite{Wu2019}. An interpretation is that the \gls{ris} coherently combines the incoming signal with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$. From the perspective of \gls{wpt}, it suggests that passive beamforming comes with a power scaling order $L^4$ under the truncated nonlinear harvester model. We then verify this observation in a simplified case where the power is uniformly allocated over multisine, all channels are frequency-flat, and $L$ is sufficiently large such that the direct channel becomes negligible. Let $X$ be the cascaded small-scale fading coefficient. The \gls{dc} in such case reduces to
			\begin{equation}
				z = \beta_2 \Lambda_{\mathrm{B}}^2 \Lambda_{\mathrm{F}}^2 \lvert X \rvert^2 L^2 P + \beta_4 \frac{2N^2 + 1}{2N} \Lambda_{\mathrm{B}}^4 \Lambda_{\mathrm{F}}^4 \lvert X \rvert^4 L^4 P^2,
			\end{equation}
			which scales quartically with $L$. Compared with active antennas, \gls{ris} elements achieve higher array gain and power scaling order, but a very large $L$ is required to compensate the double fading of the auxiliary link. These observations demonstrate the \gls{r-e} benefit of passive beamforming and emphasize the importance of accounting for the harvester nonlinearity in the waveform and beamforming design.
		\end{subsubsection}
